---
title: "Corynebacterium Diversity - Manuscript"
author: "n ahmed"
date: '2022-05-04'
output: html_document
---

###Background 

## files 
```{r}
# Sean's HOMD-classified phyloseq object
## ps_HV.PJ.eHOMD.rds 
```


## Loading libraries
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8)

library(readxl)
library(dplyr)
library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
library(ggtext)
library(RColorBrewer); packageVersion("RColorBrewer")
library(reshape2);
library(tidyverse)
library(reshape)
library(viridis); #required for the viridis color scales
library(gridExtra) #required for arraying multiple plots with grid.arrange
library(magrittr)
library(Biostrings); packageVersion("Biostrings")

#project_path <- "/Users/mahmednm/Desktop/local workspace/CoryneDiversity/bacgenR"
#path <- paste0(project_path,"/data")
```
## Load Sean's eHOMD-classified version of Payal's subsetted Phyloseq Object
```{r}
ps<-readRDS(file=file.path(project_path,"ps_HV.PJ.eHOMD.rds"))
ps
```


## Preprocessing
```{r}
#sample to an even depth
ps<-rarefy_even_depth(ps,rngseed=42,replace=FALSE)

# check for chloroplast/cynanobacteria 
unique(tax_table(ps)[,"Phylum"])
# no need to prune! 

#check for zero-sum samples 
any(sample_sums(ps) == 0) 

#  filter out spurious taxa 
#remove ASVs with < 10 reads 
psf <- prune_taxa(taxa_sums(ps) > 10, ps)
any(sample_sums(psf) == 0) 
```

## define color scheme, rename taxa  [Supplementary Figure 1]
#using sean's stable bar plot coloring
```{r}
phyToTargetLvl_classic <- function(phy) {
      #function to add a target level column to the dataframe
      #  input:  phyloseq object
      #  output: dataframe with sample metadata, taxonomy, relative abundance

      #TODO: input should be a phyloseq object and a dataframe like:
      #    mydf<-data.frame(rank=c('Phylum','Genus','Species'),
      #                   taxa=c('Bacteroidetes','Staphylococcus','Propionibacterium_acnes'),
      #                   target=c('Bacteroidetes','Staphylococcus','Cutibacterium_acnes'))
      #    mydf
      #     rank                    taxa              target
      #1  Phylum           Bacteroidetes       Bacteroidetes
      #2   Genus          Staphylococcus      Staphylococcus
      #3 Species Propionibacterium_acnes Cutibacterium_acnes
    
      joinby='tube_label' #used to join tax counts with metadata (was 'LibraryID')
      #any phylum not on this list will be collapsed into Other_Phyla
      target_phyla=c("Firmicutes","Actinobacteria","Proteobacteria","Bacteroidetes")
      #this deteremines the order of TargetLvls
      #reordered so Coryne are at bottom
      target_taxa=rev(c("C. tuberculostearicum",
                        "C. afermentans",
                        "C. accolens",
                        "Corynebacterium",
                        "Cutibacterium",
                        "Actinobacteria",
                        "Bacteroidetes",  
                        "Firmicutes",
                        "Proteobacteria", 
                        "Other"))
       #export a palette to the global environment (manually selected greens)
      corynec<-c("#A6CEE3", #ctub
                 "#1F78B4", #cafe
                 "#33A02C", #cacc
                 "#01665E") #corynebacterium
      palette<-rev(c(corynec,
                     "#9E9AC8", #Cutibacterium
                     "#5E4FA2", #Actinobacteria
                     "#F46D43", #Bacteroidetes
                     "#FDAE61", #Firmicutes
                     "#FEE08B", #proteobacteria
                     "#FFFFBF")) #other
      names(palette)<-target_taxa
      assign("classic_palette", palette, envir = .GlobalEnv)
      
      
    #TODO this should be an option
      ra = transform_sample_counts(phy, function(x) x/sum(x))

      ### This bit is Diana's for extracting a dataframe to operate on
      ### 1/21/2021 added NArm=FALSE after issues plotting data from Mihalis
      domain.phy <- tax_glom(ra, taxrank="Species",NArm = FALSE)
      tax.count <- data.frame(data.frame(domain.phy@tax_table@.Data, t(otu_table(domain.phy))))
      dfm = melt(tax.count, colnames(tax_table(domain.phy))) 

      #special case to remove leading X from numerical SampleIDs
      #dfm$variable<-substring(dfm$variable, 2)
      dfm$variable<-sub("^X","",dfm$variable,perl=TRUE)

      colnames(dfm)[colnames(dfm) == 'variable'] <- joinby
      df = plyr::join(dfm, data.frame(sample_data(domain.phy)))
      
      #now we need to add a target level column to the dataframe, default to phylum
      df$TargetLvl<-as.factor(df$Phylum) #added factor conversion 2021Aug11 (see notebook)
      #First collapse minor phyla-   other and get rid of the levels
      levels(df$TargetLvl) <- c(levels(df$TargetLvl),"Other")
      df$TargetLvl[! (df$Phylum %in% target_phyla)]<- 'Other'
      df$TargetLvl<-droplevels(df$TargetLvl)

      #add the rest of the levels to the TargetLvl factor
      levels(df$TargetLvl) <- c(levels(df$TargetLvl), setdiff(target_taxa,target_phyla))

      #These commands manually set the mapping from taxa to TargetLvl
      df$TargetLvl[df$Genus == 'Corynebacterium'] <- 'Corynebacterium'
      df$TargetLvl[df$Genus == 'Propionibacterium'] <- 'Cutibacterium'
      df$TargetLvl[df$Species == 'Corynebacterium_tuberculostearicum'] <- 'C. tuberculostearicum'
      df$TargetLvl[df$Species == 'Corynebacterium_afermentans'] <- 'C. afermentans'
      df$TargetLvl[df$Species == 'Corynebacterium_accolens'] <- 'C. accolens'

      #these are for eHOMD calls
      df$TargetLvl[df$Genus == 'Cutibacterium'] <- 'Cutibacterium'
      df$TargetLvl[df$Species == 'tuberculostearicum'] <- 'C. tuberculostearicum'
      df$TargetLvl[df$Species == 'afermentans'] <- 'C. afermentans'
      df$TargetLvl[df$Species == 'accolens'] <- 'C. accolens'

      #Reorder factors
      df$TargetLvl<-factor(df$TargetLvl, levels = target_taxa)
      
      ####  
      
      return(df)
}

#Diana Proctor wrote this to update the theme for ggplot, making barplots in a standardized fashion
#simply add this to a ggplot graph, as seen below
#customize numbers here to tweak all ggplot graphs in a uniform fashion
add_barplot_elements <- function(x) {
      theme_update() +
        theme(plot.title = element_text(color="black", size=12, face="bold.italic"),
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(size= 6, angle=90, hjust=0.5),
        axis.text.y = element_text(size = 8),
        panel.spacing = unit(0.25, "lines"),
       # strip.background = element_rect(fill = "grey"),
        strip.placement = "outside",
       # panel.spacing = unit(.2, "lines"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black", fill = "white"),
        axis.line = element_line(colour = "black"), 
        strip.background = element_rect(color = "black", fill="white"),
        strip.text.x = element_text(color="black", face="bold", size=10))
}
      
```

## Barplot: Phylum Mean Relative Abundance [Supplementary Figure 1]
```{r}
# replace HV codes with 1-23;
df0 = phyToTargetLvl_classic(psf)
df0 <- df0 %>% mutate(HV = dense_rank(Subject_ID))

p_genus =  ggplot(df0, aes(as.factor(HV), value, fill=TargetLvl)) +
        geom_bar(stat="identity", size=0.5, position="stack") + 
        ylab("Mean Relative Abundance")+
        guides(fill=guide_legend(title="Phylum")) +
        scale_colour_manual(values=classic_palette) + 
        scale_fill_manual(labels = c('Other',
                                      'Proteobacteria',
                                      'Firmicutes',
                                      'Bacterioidetes',
                                      'Actinobacteria',
                                      'Cutibacterium',
                                      'Corynebacterium',
                                      'C. accolens',
                                      'C. afermentans',
                                      'C. tuberculostearicum'), values=classic_palette) +
        theme(panel.background = element_rect(fill="black")) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank()) +
        facet_wrap(~Site, ncol=3) +
        add_barplot_elements() + 
        theme(axis.text.x = element_text(size= 5, hjust=0.5)) +
        xlab("Healthy Volunteer ID") 
        #theme(legend.text = element_text(face = 'italic'))
  
p_genus

#ggsave(file.path(path, "Phylum_RA_230522.png"), dpi = 300)

```


## phyToCoryne: define color scheme, rename taxa (Keep other) [Fig1 A]
```{r}
phyToTargetLvl_coryne <- function(phy) {
      #function to add a target level column to the dataframe
      #  input:  phyloseq object
      #  output: dataframe with sample metadata, taxonomy, relative abundance

      #TODO: input should be a phyloseq object and a dataframe like:
      #    mydf<-data.frame(rank=c('Phylum','Genus','Species'),
      #                   taxa=c('Bacteroidetes','Staphylococcus','Propionibacterium_acnes'),
      #                   target=c('Bacteroidetes','Staphylococcus','Cutibacterium_acnes'))
      #    mydf
      #     rank                    taxa              target
      #1  Phylum           Bacteroidetes       Bacteroidetes
      #2   Genus          Staphylococcus      Staphylococcus
      #3 Species Propionibacterium_acnes Cutibacterium_acnes
    
      joinby='tube_label' #used to join tax counts with metadata (was 'LibraryID')
      #any species not on this list will be collapsed into Other
      target_species=c("tuberculostearicum","afermentans","singulare",
                       "accolens","sp.HMT207","pilbarense",
                       "afermentans:pilbarense:mucifaciens","mucifaciens",
                       "simulans","minutissimum","jeikeium","appendicis",
                       "propinquum","accolens:macginleyi:tuberculostearicum",
                       "urealyticum")
      
      #this deteremines the order of TargetLvls
      #reordered so Coryne are at bottom
      target_taxa=rev(c("C. tuberculostearicum","C. afermentans","C. singulare",
                       "C. accolens","C. sp. HMT 207","C. pilbarense",
                       "C. afermentans:pilbarense:mucifaciens","C. mucifaciens",
                       "C. simulans","C. minutissimum","C. jeikeium","C. appendicis",
                       "C. propinquum","C. accolens:macginleyi:tuberculostearicum",
                       "C. urealyticum", "Other Corynebacteria"))
       #export a palette to the global environment (manually selected greens)
      corynec<-c("#A6CEE3", #ctub
                 "#1F78B4", #cafe
                 "#B2DF8A", #csing
                 "#33A02C", #cacc
                 "#FB9A99",
                 "#E31A1C",
                 "#FDBF6F",
                 "#FF7F00",
                 "#CAB2D6",
                 "#6A3D9A",
                 "#FFFF99",
                 "#eaaf16",
                 "#B15928",
                 "#683c00",
                 "#394b3f") 
      palette<-rev(c(corynec,
                     "gray")) #other
      names(palette)<-target_taxa
      assign("coryne_palette", palette, envir = .GlobalEnv)
      
      ###
      target_species=rev(c())
      
    #TODO this should be an option
      ra = transform_sample_counts(phy, function(x) x/sum(x))

      ### This bit is Diana's for extracting a dataframe to operate on
      ### 1/21/2021 added NArm=FALSE after issues plotting data from Mihalis
      domain.phy <- tax_glom(ra, taxrank="Species",NArm = FALSE)
      tax.count <- data.frame(data.frame(domain.phy@tax_table@.Data, t(otu_table(domain.phy))))
      dfm = melt(tax.count, colnames(tax_table(domain.phy))) 

      #special case to remove leading X from numerical SampleIDs
      #dfm$variable<-substring(dfm$variable, 2)
      dfm$variable<-sub("^X","",dfm$variable,perl=TRUE)

      colnames(dfm)[colnames(dfm) == 'variable'] <- joinby
      df = plyr::join(dfm, data.frame(sample_data(domain.phy)))
      
      #now we need to add a target level column to the dataframe, default to phylum
      df$TargetLvl<-as.factor(df$Species) #added factor conversion 2021Aug11 (see notebook)
      #First collapse minor phyla-   other and get rid of the levels
      levels(df$TargetLvl) <- c(levels(df$TargetLvl),"Other Corynebacteria")
      df$TargetLvl[! (df$Species %in% target_species)]<- 'Other Corynebacteria'
      df$TargetLvl<-droplevels(df$TargetLvl)

      #add the rest of the levels to the TargetLvl factor
      levels(df$TargetLvl) <- c(levels(df$TargetLvl), setdiff(target_taxa,target_species))

      #These commands manually set the mapping from taxa to TargetLvl
            #these are for eHOMD calls
      df$TargetLvl[df$Species == 'tuberculostearicum'] <- 'C. tuberculostearicum'
      df$TargetLvl[df$Species == 'afermentans'] <- 'C. afermentans'
      df$TargetLvl[df$Species == "singulare"] <- 'C. singulare'
      df$TargetLvl[df$Species == 'accolens'] <- 'C. accolens'
      df$TargetLvl[df$Species == "sp.HMT207"] <- 'C. sp. HMT 207' #prev C. imitans
      df$TargetLvl[df$Species == "pilbarense"] <- 'C. pilbarense'
      df$TargetLvl[df$Species == "afermentans:pilbarense:mucifaciens"] <- 'C. afermentans:pilbarense:mucifaciens'
      df$TargetLvl[df$Species =="mucifaciens"] <- 'C. mucifaciens'
      df$TargetLvl[df$Species == "simulans"] <- 'C. simulans'
      df$TargetLvl[df$Species == "minutissimum"] <- 'C. minutissimum'
      df$TargetLvl[df$Species == "jeikeium"] <- 'C. jeikeium'
      df$TargetLvl[df$Species == "appendicis"] <- 'C. appendicis'
      df$TargetLvl[df$Species == "propinquum"] <- 'C. propinquum'
      df$TargetLvl[df$Species == "accolens:macginleyi:tuberculostearicum"] <- 'C. accolens:macginleyi:tuberculostearicum'
      df$TargetLvl[df$Species == "urealyticum"] <- 'C. urealyticum'
 

      #Reorder factors
      df$TargetLvl<-factor(df$TargetLvl, levels = target_taxa)
      
      ####  
      
      return(df)
}
```


## Barplot: Species (Keep other) [Fig1 A]
```{r}
#subset to coryne
psc <- subset_taxa(psf, Genus=="Corynebacterium", na.rm=TRUE)
#prune samples w/o taxa 
psc <- prune_samples(sample_sums(psc) > 0, psc)
csum <- tapply(taxa_sums(psc), tax_table(psc)[, "Species"], sum, na.rm=TRUE)
csort <- names(sort(csum, TRUE))[1:15]

# replace HV codes with 1-23;
df1 = phyToTargetLvl_coryne(psc)
df1 <- df1 %>% mutate(HV = dense_rank(Subject_ID))

p_coryne =  ggplot(df1, aes(as.factor(HV), value, fill=TargetLvl)) +
        geom_bar(stat="identity", size=0.5, position="stack") +
        ylab("Mean Relative Abundance")+  xlab("")  +
        guides(fill=guide_legend(title="Species")) +
        scale_colour_manual(values=coryne_palette) +
        scale_fill_manual(values=coryne_palette) +
        theme(panel.background = element_rect(fill="black")) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank()) +
        facet_wrap(~Site, ncol=3) +
        add_barplot_elements() +
        theme(axis.text.x = element_text(size= 5, hjust=0.5)) +
        xlab("Healthy Volunteer ID")
p_coryne

#ggsave(file.path(path, "Species_RA_230622.png"), dpi = 300, width=8, height=7)
```
## ASV-level analysis [Figure 1B]
```{r}

##preserve ASV relative abundances across rows 
round_preserve_sum <- function(x, digits = 0) {
  up <- 10 ^ digits
  x <- x * up
  y <- floor(x)
  indices <- tail(order(x-y), round(sum(x)) - sum(y))
  y[indices] <- y[indices] + 1
  y / up
}

#subset species-level dataframe (psc) to just c. tuberculostearicum
dftub <- subset_taxa(psc, (tax_table(psc)[,'Species']=='tuberculostearicum'), na.rm=T)
#remove NAs
dftub <- prune_samples(sample_sums(dftub) > 0, dftub)
#count ASVs
taxa_sums(dftub)

#change Ea from 'ear' to 'sebaceous'
sample_data(dftub)$Habitat[sample_data(dftub)$Habitat %in% 'ear'] <- 'sebaceous'

#change taxa counts to relative abundance 
dftub.ra <- transform_sample_counts(dftub, function(OTU) OTU/sum(OTU))
#group taxa by habitat
dftub.ra.habitat <- merge_samples(dftub.ra, "Habitat")
#calculate relative abundance of taxa within each habitat 
dftub.ra.habitat.ra <- transform_sample_counts(dftub.ra.habitat, function(OTU) OTU/sum(OTU))

##calculating  coryne per habitat (as a fraction of all bacteria)
#collapse all-bacteria df to genus level taxa assignments 
psf.glom <- tax_glom(psf, taxrank = 'Genus', NArm = TRUE)
#change Ea from 'ear' to 'sebaceous'
sample_data(psf.glom)$Habitat[sample_data(psf.glom)$Habitat %in% 'ear'] <- 'sebaceous'
#calculate relative abundances for taxa 
df.ra <- transform_sample_counts(psf.glom, function(OTU) OTU/sum(OTU))
#calculate abundances by habitat 
df.ra.habitat <- merge_samples(df.ra, "Habitat")
df.ra.habitat.ra <- transform_sample_counts(df.ra.habitat, function(OTU) OTU/sum(OTU))
#extract c. tuberculostearicum (glommed to ASV3) RAs
c.hab <- round_preserve_sum(otu_table(df.ra.habitat.ra)[,'ASV3'],digits=2)
c.hab
#          ASV3
# dry       0.15
# feet      0.35
# moist     0.20
# nare      0.34
# sebaceous 0.04


#mean rel abundance across the 6 habitats 
sort(round(taxa_sums(dftub.ra.habitat.ra)/5,digits=3), decreasing = T)
# ASV3 constitutes 83.2% of all c.tuberculostearicum reads

#top 5 most abundant OTUs (>=0.3% mean RA; all others <=0.1% mean RA)
target_asv <- rev(c("ASV3",
                    "ASV13",
                    "ASV39",
                    "ASV30",
                    "ASV88",
                    'Other'))
 
##asv otu_table to dataframe
  ra.habitat <- t(as.data.frame(otu_table(dftub.ra.habitat.ra)))#as.data.frame(otu_table(dftub.ra.habitat.ra))
  ra.habitat <- as.data.frame(ra.habitat)
  ra.habitat$OTU <- rownames(ra.habitat)
  ra.habitat <- relocate(ra.habitat, OTU, .before = dry)
  ra.habitat <- ra.habitat %>% pivot_longer(!OTU, names_to = 'habitat', values_to = '.value')
  colnames(ra.habitat)[c(2,3)] <- c("Habitat","Abundance")
  ra.habitat


  
##classify >.3% reads as other
  asv <- c("ASV3",
                    "ASV13",
                    "ASV39",
                    "ASV30",
                    "ASV88")
  ra.habitat2 <- ra.habitat
  ra.habitat2$OTU[! (ra.habitat2$OTU %in% asv)]<- 'Other'
  #ra.habitat2$Abundance <- round_preserve_sum(ra.habitat2$Abundance, digits=2)

  
##aggregate 'other' OTUs 
rahab <- ra.habitat2[26:95,]
rahab <- aggregate(Abundance ~ OTU + Habitat, FUN = sum, data=rahab)

#just top 5
ra.habitat3 <- ra.habitat2[1:25,] 
#ra.habitat3$Abundance <- round_preserve_sum(ra.habitat3$Abundance, digits=3)
  
#recombine 'other' + 'top 5' for final df 
ra.hab <- full_join(ra.habitat3, rahab)

# round to 2 decimals, maintain sum == 1.00
ra.hab$Abundance[ra.hab$Habitat == 'moist'] <- round_preserve_sum(ra.hab$Abundance[ra.hab$Habitat == 'moist'], digits=2)
ra.hab$Abundance[ra.hab$Habitat == 'sebaceous'] <- round_preserve_sum(ra.hab$Abundance[ra.hab$Habitat == 'sebaceous'], digits=2)
ra.hab$Abundance[ra.hab$Habitat == 'dry'] <- round_preserve_sum(ra.hab$Abundance[ra.hab$Habitat == 'dry'], digits=2)
ra.hab$Abundance[ra.hab$Habitat == 'nare'] <- round_preserve_sum(ra.hab$Abundance[ra.hab$Habitat == 'nare'], digits=2)
ra.hab$Abundance[ra.hab$Habitat == 'feet'] <- round_preserve_sum(ra.hab$Abundance[ra.hab$Habitat == 'feet'], digits=2)
  
##heatmap of ASV relative abundance by habitat (Mean Relative Abundance)
  p_heat <- ggplot(ra.hab, mapping=aes(x=OTU, y=Habitat, fill = Abundance)) +
    geom_tile(color = "white", lwd = 1.5, linetype = 1) +
    geom_text(aes(label = sprintf("%0.2f", round(Abundance, digits = 2))),
              color = "black", size = 4) +
    coord_fixed() + 
 #   guides(fill = guide_colourbar(barwidth = 0.5, barheight = 25, 
  #                                label=FALSE, title=NULL)) +
    scale_fill_gradient2(low='white', mid='#21918c', high='#31688e',
                           midpoint=0.5, limits = c(0, 1)) +
    labs(y = NULL, x = expression(bolditalic('C. t.')~bold(OTU))) +
      add_barplot_elements() + 
      theme(axis.text.x = element_text(color="black",size= 10, angle=0, hjust=0.5),
            axis.text.y = element_text(color="black",size= 10, hjust=0.5, angle=90),
            axis.title.x = element_text(vjust=-1),
            axis.ticks = element_blank(),
            plot.margin = unit(c(1, 1, 0, 1), "lines"),
            legend.position = "none")  
  p_heat

#ggsave(file.path(path, "ASV_RA_170522.png"), dpi = 300)


  
```
## ctub asvs: heatmap + barplot [Figure 1B]
```{r}

###compare reads: coryne vs all bacteria 

# get ctub reads 
psf.merge1 <- tax_glom(psf, 'Species')
#change Ea from 'ear' to 'sebaceous'
sample_data(psf.merge1)$Habitat[sample_data(psf.merge1)$Habitat %in% 'ear'] <- 'sebaceous'

psf.merge1 <- merge_samples(psf.merge1, 'Habitat')
otu_table(psf.merge1)[,'ASV3'] # = sums of c. tuberculostearicum reads 
# get coryne reads vs other bacteria 
psf.merge2 <- tax_glom(psf, 'Genus')
#change Ea from 'ear' to 'sebaceous'
sample_data(psf.merge2)$Habitat[sample_data(psf.merge2)$Habitat %in% 'ear'] <- 'sebaceous'

psf.merge2 <- merge_samples(psf.merge2, "Habitat")
df.genus2 <- as.data.frame(otu_table(psf.merge2))
df.genus2 <- df.genus2 %>% relocate(ASV3, .before = ASV1)
df.genus2 <- df.genus2 %>% mutate(sum_reads=rowSums(.[2:84])) #aggregate non- c. tub
genus.sums2 <- df.genus2[,c('ASV3','sum_reads')] # = sums of corynebacteria reads 
# combine ctub reads with coryne reads vs other 
newcol <- as.data.frame(otu_table(psf.merge1)[,'ASV3']) # c. tub reads
genus.sums2 <- genus.sums2 %>% mutate('C. tuberculostearicum'=newcol$ASV3) # add c. tub reads to df 
genus.sums2$Corynebacteria <- genus.sums2$ASV3 - genus.sums2$`C. tuberculostearicum` # = tot coryne - c. tub
genus.sums2 <- genus.sums2[,-1] # remove tot coryne column 
genus.ra <- data.frame(C_tuberculostearicum=genus.sums2$`C. tuberculostearicum`/genus.sums2$sum_reads, #calculate ra
                   Corynebacteria=genus.sums2$Corynebacteria/genus.sums2$sum_reads, Other=(genus.sums2$sum_reads-df.genus2$ASV3)/genus.sums2$sum_reads)
#genus.ra$Other <- ((1 - genus.ra$C_tuberculostearicum) - genus.ra$Corynebacteria)
rownames(genus.ra) = rownames(genus.sums2)
#genus.ra[,1:3] <- round(genus.ra[,1:3], 2)



########### barplot: corynebacteria (subset: c. tuberculostearicum) vs other 
genus.ra$habitat <- rownames(genus.ra)
genus.ra <- genus.ra %>% pivot_longer(!habitat,names_to = 'group', values_to = 'value')
genus.ra$group <- factor(genus.ra$group, levels=rev(c("C_tuberculostearicum", 'Corynebacteria','Other')))
genus.ra$habitat <- factor(genus.ra$habitat, levels=rev(c('sebaceous', 'nare', 'moist', 'feet', 'dry')))


## horizontal plot to align w heatmap  
sidep <- ggplot(genus.ra, aes(habitat, value, fill=group)) +
  geom_bar(stat="identity", size=0.5, position="stack", colour="black") +
  ylab("% Bacterial Reads")+  xlab("")  + guides(fill=guide_legend(title="")) +
  scale_colour_manual(values=c('white',"#6c8491", "#6c8491"), 
                            label=c('Other','Corynebacteria','C. tuberculostearicum')) +
        scale_fill_manual(values=c('white','#6c8491', "#6c8491"),
                          label=c('Other','Corynebacteria','C. tuberculostearicum'))+
        theme(panel.background = element_rect(fill="black"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank()) +
  add_barplot_elements() +
  geom_bar_pattern(stat='identity', 
                         pattern=rev(c('none','none','stripe',
                                   'none','none','stripe',
                                   'none','none','stripe',
                                   'none','none','stripe',
                                   'none','none','stripe')),
                         pattern_fill = '#e2eef6',
                         pattern_color = '#e2eef6',
                         pattern_spacing = .01,
                         pattern_angle = 90,
                         pattern_linetype = 0.8,
                         pattern_alpha = 0.75,
                         aes(fill=group)
                         ) +
  guides(fill=guide_legend(override.aes = list(pattern=c('none','none','stripe'))))+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(color="black",size= 10, angle=0),
       # axis.ticks.length.x  = unit(0, "pt"),
        axis.ticks.y =element_blank(), 
              legend.text = element_text(size=10),
              legend.title = element_blank(),
        axis.title.x = element_text(vjust=-1),
        legend.position = 'bottom', legend.direction="vertical",
       plot.margin = unit(c(1, 0.5, 1, 1), "lines")) +
  #scale_x_discrete(expand=c(0,0))+
 scale_y_reverse(limits = c(1, 0),
                 labels= scales::percent,
                 expand = c(0,0)) + coord_flip()
        
  
# combine
sidep + p_heat + plot_layout(widths = c(.5, 2), heights=c(1,1)) 

#ggsave(file.path(path, "genus_sidebar3_170522.png"), dpi = 300, width = 12, height = 9)




```
## asv-level analysis (by subject) [incl. Supplementary Figure 2]
```{r}
dftub.ra ## c. tuberculostearicum, relative abundance transformed
sample_data(dftub.ra)
otu_table(dftub.ra)

asvs <- cbind(sample_data(dftub.ra), otu_table(dftub.ra))
#subset cols to site, subj, asvs 
asvs <- asvs[,c(3,5,11:28)]


df.asvs <- asvs %>% pivot_longer(!c(Subject_ID, Site), names_to = 'OTU', values_to = 'value')
df.asvs <- df.asvs %>% mutate(HV = dense_rank(Subject_ID)) #HV coding


####### supplementary figure 2: asv abundance by individual
#top 5 most abundant OTUs (>=0.3% mean RA; all others <=0.1% mean RA)
target_asv <- rev(c("ASV3",
                    "ASV13",
                    "ASV39",
                    "ASV30",
                    "ASV88",
                    'Other'))

##classify >.3% reads as other
  asv <- c("ASV3",
                    "ASV13",
                    "ASV39",
                    "ASV30",
                    "ASV88")
  df.asvs$OTU[! (df.asvs$OTU %in% asv)]<- 'Other'
  
df.asvs$OTU <- factor(df.asvs$OTU, levels=target_asv)

p_asv = ggplot(df.asvs, aes(as.factor(HV), value, fill=OTU)) +
        geom_bar(stat="identity", size=0.5, position="stack") +
        ylab("Mean Relative Abundance")+  xlab("")  +
        guides(fill=guide_legend(title="OTU")) +
        scale_colour_manual(values=c('gray','#fac228','#f57d15','#d44842','#801f6c','#1b0c41'),label=target_asv) +
        scale_fill_manual(values=c('gray','#fac228','#f57d15','#d44842','#801f6c','#1b0c41'),label=target_asv)+
        theme(panel.background = element_rect(fill="darkblue")) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank()) +
        facet_wrap(~Site, ncol=3) +
        add_barplot_elements() +
        theme(axis.text.x = element_text(size= 5, hjust=0.5)) +
        xlab("Healthy Volunteer ID")
p_asv

#ggsave(file.path(path, "asv_bar_240622.png"), dpi = 300, width=8, height=7)

```

## abundance related metrics for manuscript text
```{r}

##### calculating ASV abundance/habitat

#dftub.ra = c. tuberculostearicum phy object, with ASV relative abundance transformed
count(sample_data(dftub.ra)[,'Site'], Site) ## 22 Ac samples
acs <- subset(df.asvs,df.asvs$Site=='Ac') %>% group_by(OTU) 
aggregate(value~ OTU, data=acs[,3:4],FUN=sum)
## %ASV13 in Ac = 0.228164/22 = 0.01267578 -> 1.0% 

count(sample_data(dftub.ra)[,'Site'], Site) ## 21 Ph samples
phs <- subset(df.asvs,df.asvs$Site=='Ph') %>% group_by(OTU) 
aggregate(value~ OTU, data=phs[,3:4],FUN=sum)
## %ASV13 in Tw = 1.6599713/21 = 0.07904625 -> 7.9%

count(sample_data(dftub.ra)[,'Site'], Site) ## 23 Tw samples
tws <- subset(df.asvs,df.asvs$Site=='Tw') %>% group_by(OTU) 
aggregate(value~ OTU, data=tws[,3:4],FUN=sum)
## %ASV13 in Tw = 2.0866383/23 = 0.0907234 -> 9.1%

count(sample_data(dftub.ra)[,'Site'], Site) ## 21 Tn samples
tns <- subset(df.asvs,df.asvs$Site=='Tn') %>% group_by(OTU) 
aggregate(value~ OTU, data=tns[,3:4],FUN=sum)
## %ASV13 in Tw = 14.5959714/21 = 0.6950463 -> 70.0%

dftub
# 293 samples

count(unique(sample_data(dftub)[,'Subject_ID']))
# 23 subjects 

sum(otu_table(dftub)[,'ASV3'] == 0)
# 39 samples do not contain ASV3; 
# 254 samples contain ASV3

asv3s <- as.data.frame(otu_table(dftub)[,'ASV3'])
ctub_merge <- cbind(sample_data(dftub), asv3s)
ctub_merge <- ctub_merge[,c('Subject_ID', 'ASV3', 'Site')]
ctub_merge %>% group_by(Subject_ID) %>% tally()
# ASV3 is present in all subjects



count(sample_data(dftub)[,'Site'], Site)
abs %>% group_by(Site) %>% tally()

count(subset(ctub_merge, ASV3 == 0))


ctub_merge %>% group_by(ASV3) %>% tally()
pres <- subset(ctub_merge, ASV3 > 0)
abs <- subset(ctub_merge, ASV3 == 0)
unique(pres[,'Subject_ID'])
unique(abs[,'Subject_ID'])

ct_merge <- cbind(sample_data(dftub), otu_table(dftub))

```