---
title: "Heap's Law Model [Figure 3B]"
author: "Nashwa Ahmed"
date: "2023-08-11"
output: html_document
---

# Introduction

 Generate gene rarefaction curve for C. tuberculostearicum genome sequences, from which we estimate Heap's Law parameters. 

## Input 

* gene_presence_absence.Rtab - panaroo gene presence/absence table for 28 C. tuberculostearicum genomes

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

path <- getwd()
```

## Load Libraries
```{r}
library(vegan)
library(igraph)
```


## Gene rarefaction curve w/ Heap's law fit for 28 C.t. genomes (including 5 RefSeq) 
```{r}

#Read in presence/absence table 
pa <- t(read.table(file=paste0(path,'/gene_presence_absence.Rtab'),sep="\t",row.names=1,header=T))
sp <- specaccum(pa, 'random', permutations=1000)

#Fit data to Heaps' Law function, n=kN^gamma where 

  # n = richness
  # k = param a 
  # N = sites 
  # gamma = param b

fitHeaps <- nls(richness ~ a*(sites^b), sp[3:4], start=list(a=100, b=-1))

summary(fitHeaps)
#b (gamma) = 0.2976  
#a (k) = 1.977e+03

model <- predict(fitHeaps, sp[3:4])

## plot curve with Heap's law fit 

plot(sp, ci.type='poly', col='orange', lwd=4, ci.lty=0, ci.col='#FED8B1', xlab='28 C. tuberculostearicum strains', ylab='Number of genes', main='', font.lab=2)
lines(model,col='black', lwd=2)
legend(15, 1500, legend=c(as.expression(bold("Heap's Law:")~"n = k"~N^~gamma),
                          as.expression("n = 1977 *"~N^~"0.30 ± 0.01")),
       text.width = 10,
       cex = 1.2,
       bg='lightblue')
legend(15, 2950, legend=c('Pangenome size', "Heap's law fit"), cex = 1.2, col=c('orange','black'), pch='-', pt.cex = 4,text.width = 10)


```

Heap's law model for C. tuberculostearicum (strict)
```{r}
pa_strict <- pa[c(2,4:6,10:11,13,15:27),] #subset to CTC (strict) 

table(colSums(pa_strict != 0) == 0) # remove the 947/5451 genes absent from strict subset 

pa_strict <- data.frame(pa_strict) %>%
    select(where(~ any(. != 0))) %>% as.matrix()

#calculate gene accumulation curve: 20 C. t. including 5 RefSeq genomes 

sp_strict <- specaccum(pa_strict, 'random', permutations=1000)

#fit heaps law to curve

# n = richness
# k = a
# N = sites 
# gamma = b
fitHeaps_strict <- nls(richness ~ a*(sites^b), sp_strict[3:4], start=list(a=100, b=-1))
summary(fitHeaps_strict)
#b (gamma) = 0.2537  
#a (K) = 2.062e+03

model_strict <- predict(fitHeaps_strict, sp_strict[3:4])

## plot curve with Heap's law fit 
plot(sp_strict, ci.type='poly', col='#5fa2ce', lwd=4, ci.lty=0, ci.col='#a3cce9', xlab='20 C. tuberculostearicum strains', ylab='Number of genes', main='', font.lab=2)
lines(model_strict,col='black', lwd=2)
legend(10, 1500, legend=c(as.expression(bold("Heap's Law:")~"n = k"~N^~gamma),
                          as.expression("n = 2062 *"~N^~"0.25 ± 0.01")),
       text.width = 8,
       cex = 1.2,
       bg='lightblue')
legend(10, 2950, legend=c('Pangenome size', "Heap's law fit"), cex = 1.2, col=c('orange','black'), pch='-', pt.cex = 4,text.width = 8)



```

Combined plot [Figure 3B]
```{r}

#postscript(file="/3B.eps", width=6,height=4)

plot(sp, ci.type='poly', col='#57606c', lwd=4, ci.lty=0, ci.col='#a3acb9', xlab='Number of C. tuberculostearicum strains', main='', font.lab=2, las=1, ylab='')
lines(sp_strict, ci.type='poly', col='#5fa2ce', lwd=4, ci.lty=0, ci.col='#a3cce9', main='', font.lab=2)
lines(model,col='black', lwd=2)
lines(model_strict,col='black', lwd=2)
text(25,4450,as.expression(gamma~"= 0.30 ± 0.01"),cex=6/7)
text(17,3450,as.expression(gamma[strict]~"= 0.25 ± 0.01"),cex=6/7)
title( ylab=substitute(paste(bold('Number of genes'))),line = 3.1)

#dev.off()

```

Combined with colored text 
```{r}
plot(sp, ci.type='poly', col='#57606c', lwd=4, ci.lty=0, ci.col='#a3acb9', xlab='Number of C. tuberculostearicum strains', ylab='Number of genes', main='', font.lab=2)
lines(sp_strict, ci.type='poly', col='#5fa2ce', lwd=4, ci.lty=0, ci.col='#a3cce9', main='', font.lab=2)
lines(model,col='black', lwd=2)
lines(model_strict,col='black', lwd=2)
text(25,4450,as.expression(gamma~"= 0.30 ± 0.01"),cex=6/7)
text(17,3450,as.expression(gamma~"= 0.25 ± 0.01"),cex=6/7, col="#1f77b4")
```
