---
title: "Growth Curve Analysis"
author: "Nashwa Ahmed"
date: '2022-07-13'
output: html_document
---

# Introduction

 Calculate area-under-the-curve (AUC) for C. tuberculostearicum differential growth experiments.  

## Input 

* combined_raw.txt - combined growth curve data for experimental replicates (N=2)

## Setup
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8)
path <- getwd()
```

## Load libraries 

```{r}
library(ggplot2)
library(tidyverse)
library("RColorBrewer")
library(hms)
library(growthcurver)
library(ggpubr)
library(glue)
library(Hmisc)
```

## Read in combined experimental data (blank corrected)

```{r}

rfile <- paste0(path,'/combined_raw.txt')
rtable <- read.table(rfile, header = TRUE, sep = "\t", stringsAsFactors = FALSE)

# convert time format to hours 
colnames(rtable)[1] <- 'time'
rtable$time <- sapply(strsplit(rtable$time,":"),function(x) sum(as.numeric(x)*c(1,1/60,1/3600)))

```



## Map strain name, ribotype, media type, and experiment number to growth data
```{r}

# calculate growth statistics using GrowthCurver
gc_out_r <- SummarizeGrowthByPlate(rtable, t_trim=20, bg_correct = 'none')
head(gc_out_r) 

r.names <- t(as.data.frame(strsplit(gc_out_r$sample,'')))
r.names <- as.data.frame(r.names)
gcr <- gc_out_r
gcr$row <- r.names$V1
gcr$c <- r.names$V2
gcr <- gcr %>% mutate(media = case_when( 
    c %in% c("1","2", "3") ~ "BHI + Tween80",
    c %in% c("4","5", "6") ~ "Artificial eccrine sweat",
    c %in% c("7","8", "9") ~ "Artificial apocrine sweat"),
          strain = case_when( ## add strain 
    row %in% c("A","I") ~ "acroz",
    row %in% c("B","J") ~ "acvta",
    row %in% c("C","K") ~ "acvsz",
    row %in% c("D","L") ~ "acvsp",
    row %in% c("E","M") ~ "acrow",
    row %in% c("F","N") ~ "acrpm",
    row %in% c("G","O") ~ "acvtb",
    row %in% c("H","P") ~ "acvth"),
          phy = case_when(
    strain == "acrow"~ "A",
    strain == "acrpm"~ "A",
    strain == "acvtb"~ "A",
    strain == "acvth"~ "A",
    strain == "acroz"~ "B",
    strain == "acvta"~ "B",
    strain == "acvsz"~ "B",
    strain == "acvsp"~ "B"),
        exp = case_when(
    row %in% c("A","B","C", "D","E","F","G","H") ~ "1",
    row %in% c("I","J","K","L","M","N","O", "P") ~ "2"),
        variable = case_when(
         phy == "A" & media == "BHI + Tween80" ~ "1",
         phy == "A" & media == "Artificial eccrine sweat" ~ "2",
         phy == "A" & media == "Artificial apocrine sweat" ~ "3",
         phy == "B" & media == "BHI + Tween80" ~ "1",
         phy == "B" & media == "Artificial eccrine sweat" ~ "2",
         phy == "B" & media == "Artificial apocrine sweat" ~ "3"
        )
          )

gcr$phy <- factor(gcr$phy, levels=c('A','B'))
gcr$exp <- factor(gcr$exp, levels=c('1','2'))
gcr$media <- factor(gcr$media, levels=c('BHI + Tween80','Artificial eccrine sweat','Artificial apocrine sweat'))

gcr_auc <- gcr[,c('strain','phy','exp', 'media','variable','auc_e')]
```

## Perform ANOVA with post-hoc Tukey significance testing for multi-variate (ribotype, media) experiments 
```{r}

## anova: experiment + phylotype/strain (nested) + phylotype:media 
interaction <- aov(auc_e ~ exp + phy/strain + phy:media, data = gcr_auc)

summary(interaction)
#              Df Sum Sq Mean Sq F value Pr(>F)    
# exp           1   20.0    20.0   5.571 0.0198 *  
# phy           1  445.2   445.2 123.817 <2e-16 ***
# phy:strain    6   34.7     5.8   1.610 0.1495    
# phy:media     4  845.7   211.4  58.800 <2e-16 ***

interaction.tuk <- TukeyHSD(interaction)
interaction.tuk

#$`phy:media`
#                                                              diff        lwr        upr     p adj
#A:Artificial eccrine sweat-A:BHI + Tween80              -5.8286354  -8.075172 -3.5820993 0.0000000
#B:Artificial eccrine sweat-A:Artificial eccrine sweat   -3.2888177  -5.535354 -1.0422816 0.0010612
#B:Artificial apocrine sweat-A:Artificial apocrine sweat -3.6743125  -5.920849 -1.4277764 0.0002145


## make DF for plotting
  

  media <- c('BHI + Tween80', 'Artificial eccrine sweat', 'Artificial apocrine sweat')
  .y. <- c('auc_e','auc_e','auc_e')
  group1 <- c('A','A','A')
  group2 <- c('B','B','B')
  term1 <- c('B:BHI + Tween80','B:Artificial eccrine sweat','B:Artificial apocrine sweat')
  term2 <- c('A:BHI + Tween80','A:Artificial eccrine sweat','A:Artificial apocrine sweat')
  
  # get p-values 
  p.tuk <- c('0.0000098','0.0010612','0.0002145')
  p.tuk.adj <- c('1.356e-06','2.346e-10','4.470e-04')
  p.tuk.sig <- c('0.001','0.001','0.001') ## to plot 'p > 0.001' 
  y.position <- c('13.2','7.54804','8.5')
  
  p.values <- data.frame(media, .y., group1,group2,term1,term2,p.tuk,p.tuk.adj,p.tuk.sig,y.position)
  p.values$media <- factor(p.values$media,levels=c('BHI + Tween80','Artificial eccrine sweat','Artificial apocrine sweat') )
  p.values[,7:10] <- as.numeric(unlist(p.values[,7:10]))
```

## Plot AUC and statistical significance 
```{r}

## Set theme
plot_elements <- function(x) {
      theme_update() +
        theme(plot.title = element_text(color="black", size=12, face="bold.italic"),
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(size= 8, hjust=0.5),
        axis.text.y = element_text(size = 8),
        panel.spacing = unit(0.25, "lines"),
       # strip.background = element_rect(fill = "grey"),
        strip.placement = "outside",
       # panel.spacing = unit(.2, "lines"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "black", fill = "white"),
        axis.line = element_line(colour = "black"), 
        strip.background = element_rect(color = "black", fill="white"),
        strip.text.x = element_text(color="black", face="bold", size=10),
       legend.key=element_blank())
}

#Take average of replicates within experiment (4 strains/experiment)
mean.aucr <- aggregate(auc_e ~ strain + media + exp + variable, FUN = function(x) c(mean = mean(x), sd = sd(x), se = std.error(x)), data=gcr_auc)
aucr.avg <- do.call(data.frame, mean.aucr)

#Assign ribotype 
aucr.avg <- aucr.avg %>% mutate(phy = case_when(
    strain == "acrow"~ "A",
    strain == "acrpm"~ "A",
    strain == "acvtb"~ "A",
    strain == "acvth"~ "A",
    strain == "acroz"~ "B",
    strain == "acvta"~ "B",
    strain == "acvsz"~ "B",
    strain == "acvsp"~ "B")
    )


# Update media labels 
media.label <- c('BHI + Tween80' = 'BHI + 1% Tween80','Artificial eccrine sweat'='Sweat media + \n0.1% Tween80','Artificial apocrine sweat'='Sweat media + \n0.1% Tween80 + \nsynthetic lipid mixture')

aucr.avg <- aucr.avg %>% mutate(media2 = case_when( 
    media == "BHI + Tween80" ~ 'BHI + 1% Tween80' ,
    media == "Artificial eccrine sweat" ~ 'Sweat media + \n0.1% Tween80',
    TRUE ~ 'Sweat media + \n0.1% Tween80 + \nsynthetic lipid mixture')) 


aucr.avg$media <- factor(aucr.avg$media, labels=c('BHI + 1% Tween80','Sweat media + \n0.1% Tween80','Sweat media + \n0.1% Tween80 + \nsynthetic lipid mixture'))

## update media labels 
aucr.df <- aucr.avg 
levels(aucr.df$media) <- c('BHI + 1% Tween80','Sweat media + 0.1% Tween80','Sweat media + \n0.1% Tween80 + \nsynthetic lipid mixture')

## create p-values DF 
p.vals <- p.values
levels(p.vals$media) <- c('BHI + 1% Tween80','Sweat media + 0.1% Tween80','Sweat media + \n0.1% Tween80 + \nsynthetic lipid mixture')


## plot 

p.aucr2 <- ggplot(aucr.df, aes(x=phy, y=auc_e.mean)) +  
  geom_point(cex = 2, pch = 1.0, position = position_jitter(w = 0.2, h = 0), stroke=1,aes(color=exp)) +
  geom_boxplot(fill=NA, outlier.shape = NA) +
  stat_boxplot(geom = "errorbar", width = 0.15) +
  stat_summary(fun=mean, geom='point') +
  facet_wrap(~ media, labeller=labeller(media=label_wrap_gen(31), media=media.label)) + 
  plot_elements() + ylab("Average AUC") + xlab("Phylotype") + 
  guides(fill='none', color='none') + 
  scale_color_manual(values=c('#3c797d', '#995688'))

p.aucr2 + stat_pvalue_manual(tibble(p.vals), label = "<{p.tuk.sig}")

#ggsave("5A.eps", dpi = 300, width=7.5, height=5)

```


